### 2017-09-02 paper/main.tex
add latex for paper

\documentclass{acm_proc_article-sp}
\usepackage{etoolbox}
\usepackage{graphicx}
\usepackage{hanging}
\usepackage{mdframed}
\usepackage{minted}
\usepackage{tabularx}

% solution for getting around this issue to use the titlesec package:
% https://tex.stackexchange.com/questions/110546/spacing-below-section-title
\makeatletter
\newcommand\subparagraph{%
  \@startsection{subparagraph}{5}
  {\parindent}
  {3.25ex \@plus 1ex \@minus .2ex}
  {-1em}
  {\normalfont\normalsize\bfseries}}
\makeatother
\usepackage{titlesec}
\let\subparagraph\relax % You don't want to use \subparagraph

\begin{document}

\titlespacing\section{0pt}{8pt plus 4pt minus 2pt}{5pt plus 1pt minus 1pt}
\titlespacing\subsection{0pt}{5pt plus 4pt minus 2pt}{5pt plus 1pt minus 1pt}

% define code block framing style
\BeforeBeginEnvironment{minted}{
  \begin{mdframed}[skipabove=15pt,
                   skipbelow=10pt,
                   linewidth=0pt,
                   innerleftmargin=5pt,
                   innertopmargin=0pt,
                   innerbottommargin=0pt]}
\AfterEndEnvironment{minted}{\end{mdframed}}

\conferenceinfo{}{Bloomberg Data for Good Exchange 2016, NY, USA}

\title{
  Themis-ml:
}
\subtitle{
    A Fairness-aware Machine Learning Interface for \\
    End-to-end Discrimination Discovery and Mitigation
}

\numberofauthors{1}
\author{
\alignauthor
Niels Bantilan\\
       \affaddr{Arena.io}\\
       \affaddr{New York, NY}\\
       \email{niels.bantilan@gmail.com}
}

\maketitle

\begin{abstract}
As more industries integrate machine learning into socially sensitive decisions
processes like hiring, loan-approval, and parole-granting, we are at risk of
perpetuating historical and contemporary socioeconomic disparities. This is a
critical problem because on the one hand, organizations who use but do not
understand the discriminatory potential of such systems will facilitate the
widening of social disparities under the assumption that algorithms are
categorically objective. On the other hand, the responsible use of machine
learning can help us measure, understand, and mitigate the implicit historical
biases in socially sensitive data by expressing implicit decision-making mental
models in terms of explicit statistical models. In this paper we specify,
implement, and evaluate a ``fairness-aware'' machine learning interface called
themis-ml, which is intended for use by individual data scientists and
engineers, academic research teams, or larger product teams that use machine
learning in production systems.
\end{abstract}

% A category with the (minimum) three required fields
% \category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\category{H.4}{Information Systems Applications}{Miscellaneous TODO}

%\terms{TODO}

%\keywords{TODO}

\section{Introduction} In recent years, the transformative potential of machine
learning (ML) in many industries has propelled ML into the forefront of
mainstream media. From improving products and services to optimizing logistics
and operations, ML and artificial intelligence more broadly offer a wide range
of tools for organizations to enhance their internal and external capabilities.

As with any tool, we can use ML to engender great social benefit, but as
\cite{o2017weapons} emphasizes, we can also misuse it to bring about devastating
harm. In this paper, we focus on ML systems in the context of Decision Support
Systems (DSS), which are software systems that are intended to assist humans in
various decision-making contexts \cite{yoshimura2006decision,
montgomery2000evaluation, barnett1987dxplain, mysiak2005towards}. The misuse of
ML in these types of systems could potentially precipitate a widespread adverse
impact on society by introducing insidious feedback loops between biased
historical data and current decision- making \cite{o2017weapons}.

Researchers have developed many discrimination discovery and fairness-aware ML
methods \cite{kusner2017counterfactual, kamiran2012data, kamishima2012fairness,
kamiran2012decision, zemel2013learning, zafar2017fairness, dwork2012fairness,
zliobaite2015survey}, so we build on work done by others and seek to leverage
these techniques in the context of research- and product-based machine learning
applications. Our contributions in this paper are two-fold. First, we propose an
application programming interface (API) for ``Fairness-aware Machine Learning
Interfaces'' (FMLI) in the context of a simple binary classifier. Second, we
introduce themis-ml, an FMLI-compliant library, and apply it to a hypothetical
loan-granting DSS using the German Credit Dataset \cite{bache2013uci}. Our hope
is that themis-ml serves as a reference implementation that others might use and
extend for their own purposes.

%\newpage
\section{Bias and Discrimination}
Colloquially, bias is simply a preference for or against something, e.g.,
preferring for vanilla over chocolate ice cream. While this definition is
intuitive, there are two types of bias that are less obvious and important to
highlight in this paper: algorithmic and machine learning bias.

Algorithmic bias is when a set of mathematical rules favors one set of
attributes over others in relation to some target variable, like ``approve'' or
``deny'' a loan. Machine learning bias is when a trained model systematically
generates predictions that favor one group over another in relation to some
set of attributes, e.g., education, and some target variable, e.g., ``default
on credit''. While these definitions of bias are amoral, discrimination is in
essence moral, occurring when an action is based on biases resulting in the
unfair treatment of people. We define fairness as the inverse of
discrimination, meaning that a ``fairness-aware'' model is one that produces
non-discriminatory predictions.

Bias can lead to either direct (intended/explicit) or indirect
(unintended/implicit) discrimination, and the predominant legal concepts used to
determine these two types are known as disparate treatment and disparate impact
respectively \cite{barocas2016big}. As \cite{kusner2017counterfactual,
kamiran2012data} suggest, we can address disparate treatment in ML models by
simply removing all variables that are highly correlated to the protected class
of interest, in addition to the protected class itself, from the training data.
However, as \cite{kusner2017counterfactual} points out, doing so does not
necessarily mitigate discriminatory predictions and may actually introduce
unfairness into an otherwise fair system. In contrast, addressing disparate
impact is more complex because it depends on historical processes that generated
the training data, non-linear relationships between the features and protected
class, and whether we are interested in measuring individual- or group-level
discrimination \cite{dwork2012fairness}.

\section{A Fairness-aware Machine Learning Interface} So how does one measure
disparate impact and individual-/group-level discrimination in an ML-driven
product? In this section, we describe the main components of a simple
classification system, enumerate a few of the use cases that a research or
product team might have for using an FMLI, and propose an API that fulfills
these use cases.

A simple classification ML pipeline consists of five steps: data ingestion, data
preprocessing, model training, model evaluation, and prediction generation on
new e\(X_a\)mples. Data ingestion is outside the scope of this paper because it is a
highly variable process that depends on the application, often involves
considerable engineering effort, and potentially requires external stakeholder
bu\(y^{-})\)in.

Table 1 below outlines a simple classification system in terms of the core
interfaces in scikit-learn (sklearn), which is a machine learning library in the
Python programming language \cite{buitinck2013api}, and Table 2 delineates some
of the use cases that research or product teams might have to justify the use of
an FMLI.

\begin{table}
  \caption{A Simple Classification Pipeline}
  \renewcommand{\arraystretch}{1.75}
  \small\noindent\begin{tabularx}{\linewidth}{l X X}
    \textbf{API Interface} & \textbf{Function} & \textbf{E\(X_a\)mples} \\
    \hline
    \textbf{Transformer} &
      \emph{Preprocess} raw data for model training. &
      mean-unit variance scaling, min-max scaling \\
    \textbf{Estimator} &
      \emph{Train} models to perform a classification task. &
      logistic regression, random forest \\
    \textbf{Scorer} &
      \emph{Evaluate} performance of different models. &
      accuracy, f1-score, area under the curve \\
    \textbf{Predictor} &
      \emph{Predict} outcomes for new data. &
      single-classifier prediction, ensemble prediction
  \end{tabularx}
\end{table}

\section{FMLI Specification} Here we propose a high-level specification of
themis-ml, an FMLI named after the ancient Greek titaness of justice. We adopt
sklearn's principles of consistency, inspection, non-proliferation of classes,
composition, and sensible defaults \cite{buitinck2013api}, and extend them with
the following FMLI-specific principles:

\begin{hangparas}{.15in}{1}
  \textbf{Model flexibility.} Focus on fairness-aware methods that are applicable
    to a variety of model types because users might have no control or full
    control over the specific model training implementation.
\end{hangparas}
\begin{hangparas}{.15in}{1}
  \textbf{Fairness as performance.} Provide estimators and scoring metrics that
    explicitly encode a notion of both model accuracy and fairness so that models
    can optimize for both.
\end{hangparas}
\begin{hangparas}{.15in}{1}
\textbf{Transparency of fairness-utility tradeoff.} Fair models often make less
  accurate predictions [8, 13], which is an important factor when assessing
  their business impact.
\end{hangparas}

\begin{table}
  \caption{FMLI Use Cases}
  \renewcommand{\arraystretch}{1.75}
  \small\noindent\begin{tabularx}{\linewidth}{X X}
    \textbf{Use Case} & \textbf{Rationale} \\
    \hline
    Detect and reduce discrimination in a production machine learning pipeline.
      & Fairness-aware modeling aligns with team/company values,
        provides protection from legal liability. \\
    Measure individual-/group-level discrimination in data with respect to a
    protected class and outcome of interest.
      & Need to assess the potential bias resulting from training models on
        data. \\
    Preprocess raw data or post-process model predictions in a way that reduces
    discriminatory predictions generated by models.
      & Unable to change the underlying implementation of the model training
        process. \\
    Explicitly learn model parameters that produce fair predictions for a variety
    of model types.
      & Need for flexibility when experimenting with or deploying different model
        types. \\
    Evaluate the degree to which fairness-aware methods reduce discrimination
    and assess the fairness-utility tradeoff.
      & Need for assessing the business consequences or other implications of
        deploying a fairness-aware model.
  \end{tabularx}
\end{table}

\subsection{Preliminaries} In the following subsections we describe specific
methods from the ML fairness literature that map onto each of the sklearn
interfaces. Note that we only provide a high level summary of each method,
citing the original sources for more implementation details. The following
descriptions make two assumptions: (i) the positive target label \(y^{+}\)
refers to a desirable outcome, e.g. ``approve loan'', and vice versa for the
negative target label \(y^{-}\), and (ii) the protected class is a binary
variable defined as \(s \in {d, a}\), where \(X_d\)  are members of the
disadvantaged group and \(X_a\) are members of the advantaged group.

Following these conventions, we define \(X_d\), \(y^{+}\) and \(X_d\), \(y^{-}\)
as the set of observations of the disadvantaged group that are positively
labelled and negatively labelled respectively. Similarly, \(X_a\), \(y^{+}\) and
\(X_a\), \(y^{-}\)  are observations of the advantaged group that are positively
and negatively labelled respectively.

\subsection{Transformer} The main idea behind fairness-aware preprocessing is to
take a dataset D consisting of a feature set \(X_train\), target labels
\(y_train\), and protected class strain to output a modified dataset.

Massaging modifies ytrain by relabelling the target variables in such a way that
``promotes'' members of the disadvantaged protected class (e.g. ``immigrant'')
and ``demotes'' members of the advantaged class (e.g. ``citizen'')
\cite{kamiran2012data}. A ranker model R (e.g. logistic regression) is trained
on D and ranks are generated for all observations. Some of the top-ranked
observations \(X_d\), \(y^{-}\)  are ``promoted'' to \(X_d\), \(y^{+}\) and some
of the bottom-ranked observations \(X_a\), \(y^{+}\)  are ``demoted'' to
\(X_a\), \(y^{-}\) such that the proportion of \(y^{+}\) are equal in both
\(X_d\) and \(X_a\). One caveat is that this method is intrusive because it
directly manipulates \(y\) such that it no longer reflects the original value
found in the raw data.

\begin{minted}[mathescape,
               gobble=2,
               framesep=5mm,
               fontsize=\small]{python}
  from themis_ml.preprocess import Massage
  from sklearn.linear_model import LogisticRegression

  # use logistic regression as the ranking algorithm
  massager = Massage(ranker=LogisticRegression)

  # obtain a new set of labels
  y = massager.fit_transform(X, y, s)
\end{minted}

Reweighting takes a dataset D and assigns a weight to each observation using
conditional probabilities based on target labels and protected class membership
\cite{kamiran2012data}. In brief, large weights are assigned to \(X_d\),
\(y^{+}\) and \(X_a\), \(y^{-}\) , while small weights are assigned to \(X_d\),
\(y^{-}\) and \(X_a\), \(y^{+}\). The weights are then used as input to model
types that support weighted observations, which points to the main limitation of
this method, since not all classifiers can incorporate observation weights
during the learning process.

\begin{minted}[mathescape,
               gobble=2,
               framesep=5mm,
               fontsize=\small]{python}
  from themis_ml.preprocess import Reweight

  reweigher = Reweight()

  # obtain fairness-aware weights for each observation
  reweigher.fit(y, s)
  weights = reweigher.transform(y, s)
\end{minted}

Sampling is composed of two methods: the first involves uniformly sampling n
observations from each group, where n is the expected size of that group
assuming a uniform distribution. The second is to preferentially sample
observations using a ranker R, similar to the massaging method. The procedure is
to duplicate the top-ranked \(X_d\), \(y^{+}\) and \(X_a\), \(y^{-}\) while
removing top-ranked \(X_d\), \(y^{-}\) and \(X_a\), \(y^{+}\)
\cite{kamiran2012data}.

\begin{minted}[mathescape,
               gobble=2,
               framesep=5mm,
               fontsize=\small]{python}
  from themis_ml.preprocess import (
      UniformSample, PreferentialSample)
  from sklearn.linear_model import LogisticRegression

  # use logistic regression as the ranking algorithm
  uniform_sampler = UniformSample()
  preferential_sampler = PreferentialSample(
      ranker=LogisticRegression)

  # obtain a new dataset with uniform sampling
  uniform_sampler.fit(y_train, s_train)
  X, y, s = uniform_sampler.transform(X, y, s)

  # obtain a new dataset with preferential sampling
  preferential_sampler.fit(y_train, s_train)
  X, y, s = preferential_sampler.transform(X, y, s)
\end{minted}

\subsection{Estimator}

Themis-ml implements two methods for training fairness-aware models: the
prejudice remover regularizer (PRR), and the additive counterfactually fair
(ACF) model.

\cite{kamishima2012fairness} proposes PRR as an optimization technique that
extends the standard L1/L2-norm regularization method \cite{ng2004feature,
ribeiro2016should} by adding a prejudice index term to the objective function.
This term is equivalent to normalized mutual information, which measures the
degree to which predictions \(y\) and s are dependent on each other. With values
ranging from 0 to 1, 0 means that \(y\) and s are independent and a value of 1
means that they are dependent. The goal of the objective function is to find
model parameters that minimize the difference between \(y\) and \(y\) in
addition to the degree to which \(y\) depends on s.

\begin{minted}[mathescape,
               gobble=2,
               framesep=5mm,
               fontsize=\small]{python}
  from themis_ml.linear_model import LogisticRegressionPRR

  # use L2-norm regularization and prejudice index as
  # the discrimination penalizer
  lr_prr = LogisticRegressionPRR(
      penalty="L2", discrimination_penalty="PI")

  # fit the models
  lr_prr.fit(X, y, s)
\end{minted}

ACF is a method described by \cite{kusner2017counterfactual} within the
framework of counterfactual fairness. The main idea is to train linear models to
predict each feature using the protected class attribute(s) as input. We can
then compute the residuals \(\epsilon_{ij}\) between the predicted feature
values and true feature values for each observation i for each feature j. The
final model is then trained on \(\epsilon_{ij}\) as features to predict y.

\begin{minted}[mathescape,
               gobble=2,
               framesep=5mm,
               fontsize=\small]{python}
  from themis_ml.linear_model import LinearACFClassifier

  # by default, LinearACFClassifier uses linear
  # regression as the continuous feature estimator
  # and logistic regression as the binary feature
  # estimator and target variable classifier
  linear_acf = LinearACFClassifier()

  # fit the models
  linear_acf.fit(X_train, y_train, s_train)
\end{minted}

\subsection{Predictor}

Themis-ml draws on two methods to make model type-agnostic predictions: Reject
Option Classification (ROC) and Discrimination Aware Ensemble Classification
(DAEC) \cite{kamiran2012decision}. Unlike the Transformer and Estimator methods
outlined above, ROC and DAEC do not modify the training data or the training
process. Rather, they postprocess predictions in a way that reduces
discriminatory predictions.

\cite{kamiran2012decision} describes two ways of implementing ROC, starting with
ROC in a single classifier setting. ROC works by training an initial classifier
on D, generating predicted probabilities on the test set, and then computing the
proximity of each prediction to the decision boundary learned by the classifier.
Within the critical region threshold \(\theta\) around the decision boundary,
where 0.5 < \(\theta\) < 1, \(X_d\) are assigned as \(y^{+}\) and \(X_a\) are
assigned as \(y^{-}\). ROC in the multiple classifier setting is similar to the
single classifier setting, except that predicted probabilities are defined as
the weighted average of probabilities generated by each classifier.

\begin{minted}[mathescape,
               gobble=2,
               framesep=5mm,
               fontsize=\small]{python}
  from themis_ml.postprocessing import (
      SingleROClassifier, MultiROClassifier)
  from sklearn.linear_model import LogisticRegression
  from sklearn.tree import DecisionTreeClassifier

  # use logistic regression for single classifier setting
  single_roc = SingleROClassifier(
      estimator=LogisticRegression())

  # use logistic regression and decision trees for
  # multiple classifier setting
  multi_roc = MultiROClassifier(
      estimators=[LogisticRegression(),
                  DecisionTreeClassifier()])

  # fit the models and generate predictions
  single_roc.fit(X, y, s)
  multi_roc.fit(X, y, s)
  single_roc.predict(X, s)
  multi_roc.predict(X, s)
\end{minted}

The main limitation of ROC is that model types must be able to produce predicted
probabilities. DAEC gets around this problem by training an ensemble of
classifiers and, through a similar relabelling rule as ROC, re-assigns any
prediction where classifiers disagree on the predicted label. As
\cite{kamiran2012decision} notes, in general, the larger the disagreement
between classifiers, the larger the reduction in discrimination.

\begin{minted}[mathescape,
               gobble=2,
               framesep=5mm,
               fontsize=\small]{python}
  from themis_ml.postprocessing import DAEnsembleClassifier
  from sklearn.linear_model import LogisticRegression
  from sklearn.tree import DecisionTreeClassifier

  # use logistic regression and decision trees
  dae_clf = DAEnsembleClassifier(
      estimators=[LogisticRegression(),
                  DecisionTreeClassifier()])

  # fit the models and generate predictions
  dae_clf.fit(X, y, s)
  dae_clf.predict(X, s)
\end{minted}

\subsection{Scorer}

The Scorer interface is concerned with measuring the degree to which data or
predictions are discriminatory. Themis-ml implements two methods for measuring
group-level discrimination and two methods for measuring individual-level
discrimination.

In the context of measuring group-level discrimination,
\cite{zliobaite2015survey} describes mean difference and normalized mean
difference. Mean difference measures the difference between \(p(a \cup y^{+})\)
and \(p(d \cup y^{+})\). Values range from -1 to 1, where -1 is the reverse-
discrimination case (all \(X_a\) have \(y^{-}\) labels and all \(X_d\) members
have \(y^{+}\) labels) and 1 is the fully discriminatory case (all \(X_a\) have
\(y^{+}\) labels and all \(X_d\) have \(y^{-}\) labels). Normalized mean
difference scales these values to 0 and 1 based on the maximum possible
discrimination in a dataset given the rate of positive labels
\cite{zliobaite2015survey}.

\begin{minted}[mathescape,
               gobble=2,
               framesep=5mm,
               fontsize=\small]{python}
  from themis_ml.metrics import (
      mean_difference, normalized_mean_difference)

  # compare group-level discrimination in true
  # labels and predicted labels
  md_y_true = mean_difference(y, s)
  md_y_pred = mean_difference(pred, s)
  md_y_pred - md_y_true

  norm_md_y_true = norm_mean_difference(y, s)
  norm_md_y_pred = norm_mean_difference(pred, s)
  norm_md_y_pred - norm_md_y_true
\end{minted}

\cite{zliobaite2015survey} also describes consistency and situation test score
as individual-level discrimination measures. Consistency measures the difference
between the target label of a particular observation and target labels of its
neighbors. K-nearest neighbors (knn) measures the pairwise distance between
observations X. Then, for each observation \(x_i\) and each neighbor \(x_j \in
knn(x_i)\), we compute the differences between \(y_i\) and target labels of
neighbor \(y_i\). A consistency score of 0 indicates that there is no
individual-level discrimination and a score of 1 indicates that there is maximum
discrimination in the dataset.

The situation test score metric is similar to consistency, except we consider
only \(x_i \in X_d\). This method uses mean difference to compute a
discrimination score among neighbors \(x_j \in knn(x_i)\), producing a score
between 0 and 1, where 0 indicates no discrimination and 1 indicates maximum
discrimination \cite{zliobaite2015survey}.

\begin{minted}[mathescape,
               gobble=2,
               framesep=5mm,
               fontsize=\small]{python}
  from themis_ml.metrics import (
      consistency, situation_test_score)

  # compare individual-level discrimination
  # in true labels and predicted labels
  c_true = consistency(y, s)
  c_pred = consistency(y, s)
  c_pred - c_true

  sts_true = situation_test_score(y, s)
  sts_pred = situation_test_score(y, s)
  sts_pred - sts_true
\end{minted}

\section{Evaluating Themis-ml}

\emph{\textbf{Work in progress}. Using the German Credit dataset, we evaluate
the efficacy of themis-ml in reducing discriminatory predictions. We do this by
comparing the performance of various fairness-aware techniques with a standard
``fairness- unaware'' classification system, as measured by the four
discrimination measures described above.}

\section{Discussion}

In this paper, we describe FMLIs in the classification context where we consider
only a single binary protected class variable. More work needs to be done to
generalize FMLIs to the multi-classification, regression, and multiple protected
classes contexts. Additionally, little is understood about the question of the
composability of fairness-aware methods. In other words, when different
techniques are used together in sequence, are the resulting discrimination
reductions additive or otherwise?

Future technical work might also extend the FMLI specification to include
techniques like Locally Interpretable Model-Agnostic Explanations
\cite{ribeiro2016should} and develop legal frameworks for thinking about how
different stakeholders would interact with FMLIs. Continued conversations among
players in civil society, industry, academia, and government is needed to
clearly articulate which parts of an FMLI should or should not be exposed to
particular stakeholders and why. For example, the model-training components
of the FMLI should not be accessible to external auditors for intellectual
property reasons, but perhaps they might have limited access to the predictions
generated by the models.

Finally, our ability to measure and mitigate discrimination is limited by our
common social, legal, and political understanding of discrimination itself. This
common understanding is often lacking because marginalized social groups
typically do not have a voice at the table when defining what counts and does
not count as discrimination. Therefore, we see the need for continued discussion
around the meaning of discrimination in different contexts. Since FMLIs are
simply a tool to measure and mitigate formalized definitions of discrimination,
it is important for all stakeholders to engage in an inclusive forum where
everyone, especially disadvantaged social groups, can contribute.

% \begin{figure}[h]
% \centering
% \includegraphics[width=4cm]{IMG/d4gx-2015-logo.png}
% \caption{Conference logo for 2015 (PNG)}
% \label{d4gx-logo-2015}
% \end{figure}

% \begin{figure*}
%   \includegraphics[width=\textwidth]{IMG/d4gx-2015-logo.pdf}
%   \caption{Same logo, full width.}
% \end{figure*}

% \begin{figure}[h]
% \centering
% \includegraphics[width=7cm]{IMG/d4gx-2015-logo.pdf}
% \caption{Conference logo for 2015 (PDF)}
% \label{d4gx-logo-2015}
% \end{figure}

\nocite{*}
\bibliographystyle{ieeetr}
\bibliography{references}

\end{document}